### 2012-09-29 abstract.tex
shortened abstract

%% This file is to be used as a template for your submission. 
%% Rename this file and replace the text with the text of 
%% your manuscript.
%%
%% The standard LaTeX document class "article" is recommended. 
%% Use options letterpaper and 12pt.
\documentclass[letterpaper,12pt]{article}

%% This is the recommended preamble for your document.

%% Load De Gruyter specific settings 
\usepackage{dgjournal}

%% The mathptmx package is recommended for Times compatible math symbols.
%% Use mtpro2 or mathtime instead of mathptmx if you have the commercially
%% available MathTime fonts.
%% Other options are txfonts (free) or belleek (free) or TM-Math (commercial)
\usepackage{mathptmx}

% Big maths font
%\everymath{\displaystyle}
\everymath=\expandafter{\the\everymath\displaystyle}

%% Use the graphics package to include figures
\usepackage{graphicx}
\usepackage{float}

%% Use natbib with these recommended options
\usepackage[authoryear,comma,longnamesfirst,sectionbib]{natbib} 
\usepackage{url}

%% Start your document body here
\begin{document}

%% Do NOT include any fronmatter information; including the title, author names,
%% institutes, acknowledgments and title footnotes (author information, funding
%% sources, etc.). Start the document with the first section or paragraph of
%% the article.

\begin{abstract}

Tennis attracts both spectators and speculators.  Indeed, every
on-court development impacts on in-play financial markets.  For
singles matches, sophisticated traders make use of quantitative
models.  These yield the probability of each player winning the match
given the current score and the probability of each player winning a
point on serve.  Significantly, these models ignore retirement risk,
which can drastically and rapidly alter in-play odds. The precise
impact on a market depends on that market.s payout policy in the case
of player retirement.

This paper proposes novel quantitative models for the in-play
evolution of retirement risk due to injury and the corresponding
impacts on in-play match outcome markets under different retirement
payout policies. Injury occurrence is assumed to be a Bernoulli
process, with a magnitude sampled from a truncated exponential
distribution. Retirement risk is propagated from one point to another
subject to a damping parameter that reflects a player's ability to
``run off'' an injury.

An important practical challenge is to parameterise our models given a
score feed and odds data for two match outcome markets with different
retirement policies. We demonstrate the feasibility of not only
inferring model parameters from synthetic data effectively but also of
mimicking real in-play markets.

\end{abstract}

\end{document}